Architectures and Open pNets

### 2018-11-29 2018-Arch-pNets/arch-pNets.tex
very small stuff...

\newcommand{\Simon}{\\\hfill\mdash Simon}

\newcommand{\noteSB}[2][color=green!40, size=\tiny]{\todo[#1]{{#2}\Simon}}
\newcommand{\noteSBin}[2][inline,color=green!40]{\todo[#1]{{#2}\Simon}}
\newcommand{\todoSB}[2][color=green!40, size=\tiny]{\todo[#1]{\textbf{To-do Simon:} {#2}}}
\newcommand{\todoSBin}[2][inline,color=green!40]{\todo[#1]{\textbf{To-do Simon: } {#2}}}
\newcommand{\Ludo}{\\\hfill\mdash Ludo}
\newcommand{\noteLH}[2][color=orange!40, size=\tiny]{\todo[#1]{{#2}\Ludo}}
\newcommand{\noteLHin}[2][inline,color=orange!40]{\todo[#1]{{#2}\Ludo}}
\newcommand{\todoLH}[2][color=orange!40, size=\tiny]{\todo[#1]{\textbf{To-do Ludo:} {#2}}}
\newcommand{\todoLHin}[2][inline,color=orange!40]{\todo[#1]{\textbf{To-do Ludo: } {#2}}}

@@

\institute{%
  INRIA Lille -- Nord Europe, Villeneuve d'Ascq, France\\
  \email{simon.bliudze@inria.fr}
\and
\and
}

@@

\noteLHin{I believe we should simplify this paragraph about param actions}
The pNet model in \secn{pNets} relies on a notion of parameterised actions, that are
symbolic expressions using data types and variables. As we aim at encoding the low-level 
behaviour of possibly very different
programming languages, we do not want to impose one specific algebra
for denoting actions, nor any specific communication mechanism. So we
leave unspecified the constructors of the algebra that will allow building
expressions and actions. Moreover, we use a generic {\em action interaction}
mechanism, based on (some sort of) unification between two or more action
expressions, to express various kinds of communication or
synchronisation mechanisms.

@@

Formally, we assume the existence of a term algebra $\Talg$,
where $\signature$ is the signature of the data and action constructors,
and $\variables$ a set of \emph{variables}. Within $\Talg$, we distinguish a set of
\emph{data expressions} $\exprs{\variables}$ \replaceSB{, including}{and} a set of \emph{Boolean
expressions} \todoLH{$\mathcal{B} \to g$}
$\boolexprs{\variables}\subseteq\exprs{\variables}$.
On top of $\exprs{\variables}$ we build the \emph{action algebra}
$\actions{\variables}$, with $\actions{\variables}\subseteq\Talg$ \replaceSB{,}{and}
$\exprs{\variables}\cap\actions{\variables} = \emptyset$;
naturally action terms will use data expressions as subterms.
\addSB{In addition to the above notations, we will use
\[
\assigns{\variables} \bydef{=}
\Setdef{(x_i := e_i)^{i \in I}}{
  x_i^{i \in I}\subseteq \variables,
  e_i^{i \in I}\subseteq \exprs{\variables},
  I \text{ is a finite index set}
}
\]
to denote the set of \emph{variable assignments}.}  We use the symbol
$\noop$ to denote the empty assignment, \ie an assignment with
$I = \emptyset$.

To be able to reason about the data flow between pLTSs, we
distinguish \emph{input variables} of the form $?x$ within terms; the function
$\vars(t)$ identifies the set of variables in a term
$t\in\Talg$, and $iv(t)$ returns its input variables.
Action algebras can \replaceSB{encode naturally }{naturally encode }usual point-to-point message passing calculi (using 
$a(?x_1,\dots,?x_n)$ for inputs, $a(v_1,\dots,v_n)$ for outputs), but it also allows
for more general mechanisms, like gate negotiation in Lotos, or broadcast
communications.

@@

\begin{example}\emph{An open-transition.}
  \label{OT:enable-composed}
  The \texttt{EnableCompL} pNet of Fig. \ref{schema:enable-composed} has 2 controllers 
  and 2 holes. One of its possible open-transition is\footnote{Note the underlined action notation that we use to represent synchronised actions, i.e. actions that are the result of a synchronisation and that we declare as \emph{non-observable}. Those actions are only used for monitoring purposes and statement of properties but cannot be synchronised again in a synchronisation vector. This allows us to define a weak bisimilarity equivalence with p=compositional properties.}: 
\noteLH{underline notation might be a bit too much for this paper : see at the end if we remove it or not (will we use it?)}

@@

\Fig{fig:FailureTimer:pNet} shows the encoding of the Failure Monitor
architecture, including the encodings of the two coordinators, \ie
$\nopri[\NameTimer]$ and $\nopri[\NameCtrl]$.  Notice that the
encoding in the figure is slightly optimised: some of the ports do not
have an associated Boolean value, nor the additional loop transitions.
\todoSB{Revise this if necessary}\hl{We will explicitate this optimisation once we have defined the
encoding of the interaction model.}

@@

\begin{figure*}[t]
  \fbox{%
  \includegraphics[width=\textwidth]{BIPspec-ArchFailureTimerMax-v4-no-pri}}
  \caption{Encoding of the Failure Monitor architecture
    (\fig{schema:ArchFailure:BIP}) as an open pNet (without the
    priority model) \hl{\bf (Error in the picture: First two positions in the synchronisation vectors must be transposed)}}
  \label{fig:enc:nopri}
\end{figure*}

@@

The encoding of the interaction model is based on its representation
as a set of connectors. \noteSB{%
%
  A legitimate question is whether we actually need to do this,
  instead of listing explicitly the individual interactions, then
  associating a synchronisation vector to each one of them?  One of
  the factors is the encoding pNets$\rightarrow$SAT: is it better to
  have more or fewer SVs if the price is additional Boolean variables?
  The other factor might be the facility of maximal progress
  encoding\mdash TBC.  Traceability.  Anything else?
%
} Indeed, as illustrated by the Failure Monitor architecture in
\fig{schema:ArchFailure:BIP}, each connector can define several
allowed interactions, depending on its hierarchical structure and the
use of syncrhons and trigger.  \noteSB{Move this to
  \secn{components}?}For instance, the connector {\NameTimer.\PortTick
  \trigsynch (\PortFail \trigsynch \NameCtrl.\PortFail)} is a
two-level hierarchical connector.  In the second-level subconnector
{\PortFail \trigsynch \NameCtrl.\PortFail}, the port {\PortFail} is a
trigger, whereas {\NameCtrl.\PortFail} is a synchron.  Therefore, this
subconnector allows two interactions: one with {\PortFail} alone and
another one involving both ports.  Similarly, at the top level,
{\NameTimer.\PortTick} is a trigger, whereas the second-level
subconnector is a synchron.  Hence, the overall connector allows the
interaction involving only one port {\NameTimer.\PortTick}\footnote{%
%
  Such interactions are called \emph{singleton interactions}.
%
} and any interaction involving both {\NameTimer.\PortTick} and
\emph{some} interaction of the second-level subconnector.  Combining
the two levels we can see that the entire connector defines the
following three interactions (observe that $\topinterval +
\NameTimer.t = \topinterval$ and $\NameTimer.z \cap \topinterval =
\NameTimer.z$):
%
\begin{align*}
  \bigl(\{\NameTimer.\PortTick\}&,
  \true, 
  \NameTimer.z := \NameTimer.z\bigr)
  \,,\\
  \bigl(\{\PortFail, \NameTimer.\PortTick\}&,
  \true,
  \NameTimer.z := \NameTimer.z\bigr)
  \,,\\
  \bigl(\{\NameCtrl.\PortFail, \PortFail, \NameTimer.\PortTick \}&,
  \true,
  \NameTimer.z := \NameTimer.z \cap (\NameCtrl.\mathit{zone} + \NameTimer.t)\bigr)
  \,.\
\end{align*}

>>>>

\newcommand{\Simon}{\\\hfill\mdash Simon}
\newcommand{\Eric}{\\\hfill\mdash Eric}

\newcommand{\noteSB}[2][color=green!40, size=\tiny]{\todo[#1]{{#2}\Simon}}
\newcommand{\noteSBin}[2][inline,color=green!40]{\todo[#1]{{#2}\Simon}}
\newcommand{\todoSB}[2][color=green!40, size=\tiny]{\todo[#1]{\textbf{To-do Simon:} {#2}}}
\newcommand{\todoSBin}[2][inline,color=green!40]{\todo[#1]{\textbf{To-do Simon: } {#2}}}
\newcommand{\Ludo}{\\\hfill\mdash Ludo}
\newcommand{\noteLH}[2][color=orange!40, size=\tiny]{\todo[#1]{{#2}\Ludo}}
\newcommand{\noteLHin}[2][inline,color=orange!40]{\todo[#1]{{#2}\Ludo}}
\newcommand{\todoLH}[2][color=orange!40, size=\tiny]{\todo[#1]{\textbf{To-do Ludo:} {#2}}}
\newcommand{\todoLHin}[2][inline,color=orange!40]{\todo[#1]{\textbf{To-do Ludo: } {#2}}}
\newcommand{\todoEM}[2][color=green!40, size=\tiny]{\todo[#1]{\textbf{To-do Eric:} {#2}}}

@@

\institute{%
  INRIA Lille -- Nord Europe, Villeneuve d'Ascq, France\\
  \email{simon.bliudze@inria.fr}
  \and
  \and
    Universit\'e C\^ote d'Azur, Inria, CNRS, I3S, 06902 Sophia-Antipolis, France\\
  \email{eric.madelaine@inria.fr}
}

@@

\noteLHin{I believe we should simplify this paragraph about param actions

  I agree with this, but also we may look at what is useful here, and what is only used later in pLTS/pNet definitions. 
\Eric}
The pNet model in \secn{pNets} relies on a notion of parameterised actions, that are
symbolic expressions using data types and variables. As we aim at encoding the low-level 
behaviour of possibly very different
programming languages, we do not want to impose one specific algebra
for denoting actions, nor any specific communication mechanism. So we
leave unspecified the constructors of the algebra that will allow building
expressions and actions. Moreover, we use a generic {\em action interaction}
mechanism, based on (some sort of) unification between two or more action
expressions, to express various kinds of communication or
synchronisation mechanisms.

@@

Formally, we assume the existence of a term algebra $\Talg$,
where $\signature$ is the signature of the data and action constructors,
and $\variables$ a set of \emph{variables}. Within $\Talg$, we distinguish a set of
\emph{data expressions} $\exprs{\variables}$ and a set of \emph{Boolean
expressions} \todoLH{$\mathcal{B} \to g$}
$\boolexprs{\variables}\subseteq\exprs{\variables}$.
On top of $\exprs{\variables}$ we build the \emph{action algebra}
$\actions{\variables}$, with $\actions{\variables}\subseteq\Talg$ \replaceSB{,}{and}
$\exprs{\variables}\cap\actions{\variables} = \emptyset$;
naturally action terms will use data expressions as subterms.
\addSB{In addition to the above notations, we will use
\[
\assigns{\variables} \bydef{=}
\Setdef{(x_i := e_i)^{i \in I}}{
  x_i^{i \in I}\subseteq \variables,
  e_i^{i \in I}\subseteq \exprs{\variables},
  I \text{ is a finite index set}
}
\]
to denote the set of \emph{variable assignments}.}  We use the symbol
$\noop$ to denote the empty assignment, \ie an assignment with
$I = \emptyset$.

To be able to reason about the data flow between pLTSs, we
distinguish \emph{input variables} of the form $?x$ within terms; the function
$\vars(t)$ identifies the set of variables in a term
$t\in\Talg$, and $iv(t)$ returns its input variables.
Action algebras can naturally encode usual point-to-point message passing calculi (using 
$a(?x_1,\dots,?x_n)$ for inputs, $a(v_1,\dots,v_n)$ for outputs), but it also allows
for more general mechanisms, like gate negotiation in Lotos, or broadcast
communications.

@@

\begin{example}\emph{An open-transition.}
  \label{OT:enable-composed}
  \todoEM{remplacer cette OT par une du Timer, qui aurait tous les elements presents...}
  The \texttt{Failure TimerMonitor} pNet of \Fig{FailureTimer:pNet} has 2 controllers 
  and 2 holes. One of its possible open-transition is\footnote{Note the underlined action notation that we use to represent synchronised actions, i.e. actions that are the result of a synchronisation and that we declare as \emph{non-observable}. Those actions are only used for monitoring purposes and statement of properties but cannot be synchronised again in a synchronisation vector. This allows us to define a weak bisimilarity equivalence with p=compositional properties.}: 
\noteLH{underline notation might be a bit too much for this paper : see at the end if we remove it or not (will we use it?)}

@@

\Fig{FailureTimer:pNet} shows the encoding of the Failure Monitor
architecture, including the encodings of the two coordinators, \ie
$\nopri[\NameTimer]$ and $\nopri[\NameCtrl]$.  Notice that the
encoding in the figure is slightly optimised: some of the ports do not
have an associated Boolean value, nor the additional loop transitions.
\todoSB{Revise this if necessary}\hl{We will explicitate this optimisation once we have defined the
encoding of the interaction model.}

@@

%% \begin{figure*}[t]
%%   \fbox{%
%%   \includegraphics[width=\textwidth]{BIPspec-ArchFailureTimerMax-v4-no-pri}}
%%   \caption{Encoding of the Failure Monitor architecture
%%     (\fig{schema:ArchFailure:BIP}) as an open pNet (without the
%%     priority model) \hl{\bf (Error in the picture: First two positions in the synchronisation vectors must be transposed)}}
%%   \label{fig:enc:nopri}
%% \end{figure*}

@@

The encoding of the interaction model is based on its representation
as a set of connectors. \noteSB{%
%
  A legitimate question is whether we actually need to do this,
  instead of listing explicitly the individual interactions, then
  associating a synchronisation vector to each one of them?  One of
  the factors is the encoding pNets$\rightarrow$SAT: is it better to
  have more or fewer SVs if the price is additional Boolean variables?
  The other factor might be the facility of maximal progress
  encoding\mdash TBC.  Traceability.  Anything else?
%
} Indeed, as illustrated by the Failure Monitor architecture in
\fig{schema:ArchFailure:BIP}, each connector can define several
allowed interactions, depending on its hierarchical structure and the
use of synchrons and triggers.  \noteSB{Move this to
  \secn{components}?}For instance, the connector {\NameTimer.\PortTick
  \trigsynch (\PortFail \trigsynch \NameCtrl.\PortFail)} is a
two-level hierarchical connector.  In the second-level subconnector
{\PortFail \trigsynch \NameCtrl.\PortFail}, the port {\PortFail} is a
trigger, whereas {\NameCtrl.\PortFail} is a synchron.  Therefore, this
subconnector allows two interactions: one with {\PortFail} alone and
another one involving both ports.  Similarly, at the top level,
{\NameTimer.\PortTick} is a trigger, whereas the second-level
subconnector is a synchron.  Hence, the overall connector allows the
interaction involving only one port {\NameTimer.\PortTick}\footnote{%
%
  Such interactions are called \emph{singleton interactions}.
%
} and any interaction involving both {\NameTimer.\PortTick} and
\emph{some} interaction of the second-level subconnector.  Combining
the two levels we can see that the entire connector defines the
following three interactions (observe that $\topinterval +
\NameTimer.t = \topinterval$ and $\NameTimer.z \cap \topinterval =
\NameTimer.z$):
%
\begin{align*}
  \bigl(\{\NameTimer.\PortTick\}&,
  \true, 
  \NameTimer.z := \NameTimer.z\bigr)
  \,,\\
  \bigl(\{\PortFail, \NameTimer.\PortTick\}&,
  \true,
  \NameTimer.z := \NameTimer.z\bigr)
  \,,\\
  \bigl(\{\NameCtrl.\PortFail, \PortFail, \NameTimer.\PortTick \}&,
  \true,
  \NameTimer.z := \NameTimer.z \cap (\NameCtrl.\mathit{zone} + \NameTimer.t)\bigr)
  \,.\
\end{align*}
